An Open Repository of Real-Time COVID-19 Indicators

### 2021-06-15 indicators/paper/paper.tex
Updates throughout, new discussion

\dropcap{P}ublic health decision-makers, healthcare providers, epidemiological
researchers, employers, institutions, and the general public benefit from
promptly and readily accessible data regarding COVID-19 activity levels,
countermeasures, and pandemic impact.  Real-time indicators of COVID-19 activity
levels, such as statistics on cases, deaths, test positivity, and
hospitalizations, enable reports and interactive dashboard applications for
situational awareness \cite{Dong:2020, NYTimes, USAFacts}, and are essential for
most analyses of the pandemic.  These data are available for locations across
the United States from a number of official sources and independent aggregators
in varied and inconsistent formats.  Different data types and sources vary in
timeliness, based on when measured events occur in the progression of the
disease, testing capabilities, the reporting pipeline, and their publication
schedules.

@@

The Delphi Group works with partner organizations and public data sets to build
a massive database of indicators tracking COVID-19 activity and other relevant
phenomena, which has been publicly available and continuously updated since
April 2020. Alongside public data on reported cases and deaths, this database
includes several unique data streams, including indicators extracted from
de-identified medical claims data, antigen test results from a major testing
manufacturer, large-scale public surveys that measure symptoms and public
behavior, and symptom indicators based on Google search query volume. (We use
the terms ``indicator'' and ``signal'' interchangeably.) We make aggregate
signals publicly available, generally at the county level, via the COVIDcast API  
\cite{CovidcastAPI}. We store and provide access to all previous (historical)
versions of the signals, a key feature that exposes the effects of data
revisions.  Lastly, we provide R \cite{CovidcastR} and Python \cite{CovidcastPy}  
packages to facilitate interaction with the API, and an online dashboard to
visualize the data \cite{CovidcastViz}.

% RJT: We should fiddle with the bib so that the URLs actually appear in the
% references at the end.  Otherwise it's kind of pointless to cite manuals or
% websites like the COVIDcast API site or R/Python manuals.

@@

\paragraph{Internet-Based Surveys} Conducted in partnership with Facebook,
Delphi's COVID Tracking and Impact Survey receives about 50,000 responses daily,  
and has received over 20 million responses since April 2020 \cite{DelphiSurvey,
  Kreuter:2020}. From the surveys, we construct indicators on symptoms, social 
distancing, vaccination, and other attitudes and behaviors related to COVID.

@@

\begin{table*}[t]
  \centering
  \caption{Data sources available in Delphi's COVIDcast API \cite{CovidcastAPI},
    as of date of publication. The first group of data sources are produced by
    Delphi from data not otherwise available publicly (or only available in
    limited form); the second group is mirrored from public sources.}
  \begin{tabular}{>{\raggedright}p{1.2in} p{4.0in} l >{\raggedright\arraybackslash}p{0.5in}}
    \toprule
    \textbf{Data source} & \textbf{Signals available} & \textbf{First date} &
\textbf{Resolution} \\\midrule
    % Exclusive data
    Change Healthcare & Percentage of outpatient visits with COVID diagnostic
codes or codes indicating COVID-like symptoms. Based on de-identified claims
data processed by Change Healthcare. & 2020-02-01 & County$^*$ \\
    Doctor Visits & Percentage of outpatient visits primarily about COVID-like
symptoms, based on de-identified claims data provided by health system
partners. & 2020-02-01 & County$^*$ \\
    Hospital Admissions & Percentage of new hospital admissions with COVID
diagnostic codes, based on de-identified claims data provided by health system
partners. & 2020-02-01 & County$^{**}$\\
    Quidel & Test positivity rates for COVID-19 antigen tests produced by
Quidel. & 2020-05-26 & County$^{**}$ \\
    SafeGraph & Mobility metrics, such as time away from home or visits to bars 
and restaurants, based on cell phone mobility data collected by SafeGraph
\cite{SafeGraphSocial, SafeGraphPatterns}. & 2019-01-01 & County \\
    COVID Tracking and Impact Survey & COVID symptoms, social distancing
behaviors, mental health, economic impact, behavior (e.g., mask-wearing,
vaccination attitudes), and COVID testing signals based on daily surveys
conducted nationally by Delphi through Facebook \cite{DelphiSurvey,
Kreuter:2020}. & 2020-04-06 & County$^{**}$ \\
    \midrule
     % Mirrored data from other sources
    Health \& Human Services & Counts of hospital admissions due to confirmed or 
suspected COVID-19, as reported by the Department of Health \& Human Services. &
2019-12-31 & State \\
    COVID Act Now & COVID-19 testing results, such as positivity rate and number
of tests, compiled by COVID Act Now from CDC reporting. & 2020-03-02 &
County$^*$ \\
    Google Symptoms & Trends in Google search volume for terms related to
anosmia and ageusia (loss of smell or taste), which correlate with COVID
activity, based on data shared by Google \cite{GoogleSymptoms}. & 2020-02-13 &
County$^{***}$ \\
    Cases and Deaths & Confirmed COVID-19 cases and deaths, compiled by JHU CSSE 
\cite{Dong:2020} and by USAFacts \cite{USAFacts}. & 2020-01-22 & County \\
    NCHS Mortality & Weekly totals of deaths broken down by cause, such as
COVID, flu, or pneumonia, compiled by the National Center for Health Statistics
[TODO cite]. & 2020-01-26 & State \\
    \bottomrule
  \end{tabular}
  \addtabletext{$^*$Available at $>60$\% of counties; $^{**}$available at
    20--60\% of counties; $^{***}$available at $<20$\% of counties.  For some
    signals, location availability varies over time, e.g., due to variable
    reporting volume.}
  \label{tab:sources_signals}
\end{table*}

@@

Because each data source reports data in different formats, we must convert each
source to a common format. In this format, each record represents an observation
of one quantity at one time point in one location. Locations are coded
consistently using standard identifiers such as FIPS codes; the sample size and
standard error for each observation is also reported when applicable. Each
signal is reported at the finest geographic resolution its source supports (such
as county or state) and also aggregated to metropolitan statistical areas,
Health and Human Services regions, and Hospital Referral Regions. National
averages are also provided. Crucially, each record includes an \textit{issue
  date} recording when the value was first issued, as described below. This
allows tracking of multiple revisions to individual observations, each revision
with a separate issue date.

When appropriate, additional post-processing (often nontrivial) is applied to
the data. For example, data on visits to doctor's offices is subject to strong
day-of-week effects, and so regression is used to adjust for these
effects. Other indicators are available in raw versions and versions smoothed
with a seven-day moving average. All processing is done using open-source code
written primarily in Python and R, and available publicly at
\url{https://github.com/cmu-delphi/covidcast-indicators/}.

@@

Since April 2020, Delphi has been maintaining and continually improving various
online visualization tools for the COVIDcast indicators \cite{CovidcastViz}.
These tools fetch data directly from the API, and allow for exploration of both 
temporal (e.g., time series graphs) and spatial (e.g., choropleth maps) trends
in the signals, as well as many other aspects, such as correlations, anomalies,
and backfill.  There is also a dedicated dashboard for exploring results from
the COVID Tracking and Impact Survey.

@@

Many of the indicators in the COVIDcast API are intended to track COVID
activity. Five indicators in particular have the closest connections to
confirmed cases:
\begin{itemize}
\item Change Healthcare COVID-like illness (CHNG-CLI): The percentage of
  outpatient visits that are primarily about COVID-related symptoms, based on
  de-identified Change Healthcare claims data.
\item Change Healthcare COVID (CHNG-COVID): The percentage of outpatient visits
  with confirmed COVID-19, based on the same claims data.
\item COVID Tracking and Impact Survey CLI (CTIS-CLI): The estimated percentage
  of the population with COVID-like illness based on Delphi's surveys of
  Facebook users.
\item COVID Tracking and Impact Survey CLI-in-community (CTIS-CLI-in-community):
  The estimated percentage of the population who knows someone who is sick,
  based on Delphi's surveys of Facebook users.
\item Quidel COVID test positivity rate (Quidel-TPR): The percentage of positive
  test results among Quidel COVID antigen tests.
\end{itemize}
Figure~\ref{fig:time_trends_national} compares the first three of these signals
to reported COVID cases in the United States (from JHU CSSE, and smoothed with a
7-day trailing average) over the first year of the pandemic (April 15, 2020 to
April 15, 2021), illustrating how they  appear to track national trends in cases
quite well. Importantly, this same relationship persists broadly across multiple
resolutions, down to smaller geographic regions such as states and counties, as
shown in the supplement. This will also be illustrated in a more detailed
correlation analysis in the next subsection.

% RJT: @AR Note what I claimed we would do above!  I'm thinking we can just
% produce a big batch of case trends over the same time period and for the same
% signals, but for each US state (and territory?), and for the 50 most populous
% counties. AND maybe for this (below national) we should use CLI-in-community
% rather than CLI, from the survey

@@

% RJT: @AR Consider for the supplement.
% - Hospitalization trends?
% - Hospitalization correlations?
%  (I see you already have a plot for the first ... but I'm a bit bothered by
%  the fact that the HHS data looks to have such strong weekday effects that we
%  don't attempt to correct for ... perhaps we use do a 7-day rolling avg. Made
%  easy with modeltools)
% If we do so, we should refer to it in the text, this subsection (for time
% trends) and next (for correlations)

@@

To quantify the ability of the signals described above to track trends in COVID
cases, we use the Spearman (rank) correlation and analyze two key correlation
patterns, between each signal and confirmed COVID case rates (cases per 100,000
people):
\begin{enumerate}
\item \textit{Geo-wise correlations} (i.e., on a specific date, do values of the
  signal correlate with case rates across locations?): Formally, let $X_t$ and
  $Y_t$ be vectors of values of a particular signal and case rates, over all
  locations, on date $t$. The geo-wise correlation at time $t$ is defined as
  $\cor(X_t, Y_t)$. This examines whether a signal has the capability to help
  spot locations with high case rates at any given time.

@@

Figure~\ref{fig:geo_wise_correlation} shows the geo-wise correlations achieved
by the five signals and COVID case rates (from USAFacts, and again smoothed
using a 7-day trailing average), from April 15, 2020 to April 15, 2021. The
locations we consider here are counties with at least 500 cumulative cases by
the end of this period, and at which all signals are available (956 counties in
total). The  significantly positive correlations suggest that these signals 
could be useful in hotspot detection  (identifying counties with relatively high
COVID activity, at a given time). Somewhat surprisingly, the CLI-in-community
signal derived from the survey shows the strongest correlations for much of the
time period. This clearly demonstrates the value of a large-scale survey such as
CTIS for tracking symptoms and case trends, especially when other data is
unavailable.

@@

Figure~\ref{fig:time_wise_correlation} summarizes time-wise correlations from 
these five signals over the same time period, and for the same set of
counties. For each signal, we display the set correlations that it achieves in
histogram form (more precisely, using a kernel density estimate). All signals
produce positive correlations in the majority of counties considered (with very
little mass in each estimated density being to the left of zero). The largest
correlations, in bulk, are achieved by the CHNG-COVID signal; the
CTIS-CLI-in-community is a close second, and the CHNG-CLI signal is third.
There are two remarks deserving attention:

\begin{itemize}
\item This is different from what is observed in
  Figure~\ref{fig:geo_wise_correlation}, where the CTIS-CLI-in-community signal
  achieves clearly the highest correlations for most of the time
  period. However, it is worth emphasizing that time-wise and geo-wise
  correlations are truly measuring different properties of a signal; and the
  claims signals (CHNG-COVID and CHNG-CLI) seem more appropriate for
  temporal---rather than spatial---comparisons.  We revisit this point in the
  discussion.
\item It is still quite impressive (and surprising) that the
  CTIS-CLI-in-community signal, based on people reporting on the symptoms of
  others around them, can achieve  nearly as strong time-wise correlations to
  confirmed cases as can a signal that is based on picking up the occurrence of
  a confirmed case passing through the outpatient system.
\end{itemize}

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=\columnwidth]{fig/bexar_compare.pdf}
  \caption{Reported cases per day in Bexar County, Texas during the summer of
    2020. Only July 16, 4,810 backlogged cases were reported, though they
    actually occurred over the preceding two weeks (this shows up as a prolonged
    spike in the left panel due to the 7-day trailing averaging applied to the
    case counts). Daily CTIS estimates of CLI-in-community showed more stable
    underlying trends.}
  \label{fig:bexar-compare}
\end{figure}

@@

As a further example, Figure~\ref{fig:bexar-compare} displays data from Bexar
County, Texas (which contains San Antonio) during July 2020. On July 16, 2020,
4,810 backlogged cases were reported after reporting problems prevented them
from being reported over the past two weeks \cite{Palacios:2021}. This results
in a clearly visible spike in the left-hand panel of the figure (case data from
JHU CSSE, smoothed with a 7-day trailing average).  However, Delphi's COVID
Tracking and Impact Survey averaged around 350 responses per day in Bexar
County over the same time period, and was able to estimate the fraction
of the population who know someone in their community with COVID-Like Illness
(CLI). As we can see in the right-hand panel of the figure, this signal was not
affected by Bexar County's reporting problems and, as shown in the last
subsection, it is (in general) highly correlated with case rates, providing an
alternate stream of data about COVID activity unaffected by backlogs. Similar
reporting problems have occurred in many jurisdictions across the United States,
making it valuable to cross-check against external sources not part of the same
reporting systems.

@@

@@

% RJT: @AR Since we already stuff on claims, can we do: mobility, masks, vaccine
% uptake or hesitancy, distancing behaviors, how worried are you?

% How about this (in each case, each point in said plot is a county's signal
% values averaged over some period)
% - Bivariate plot of mask use versus how worried are you
% - Bivariate plot of of how worried are you vs case rates
%- Bivariate plot of of how worried are you vs hospitalizations
% I would be curious to see all three.  And perhaps color code the points by
% region to highlight regional trends

@@

% AR: I think our discussion should also make an argument: that providing many
% data sources in a common format, with revision tracking, is a valuable goal.
% Each of our individual data sources has various limitations -- SafeGraph has
% various biases, the claims data is limited by geographically varying market
% share, the surveys have demographic biases, etc. So if the argument is solely
% "We have better data than everyone else," it's a tough sell; if the argument
% is that our data being convenient to use and accessible in a single unified
% way makes it more valuable than the individual parts, and that other
% epidemiological data should be available the same way, then I think we have a
% strong case.
% RJT: I agree

% RJT Proposed format for this discussion.
% 1. COVID indicators can help supplement public health reports by filling out
% the severity pyramid, improving timeliness, helping with robustness, etc.
% 2. They also provide info well beyond public health reports (behaviors,
% attitudes)
% 3. There is an argument to be made for having data all in one unified
% convenient access format (with all revisions)

% There are a number of open questions, and challenges to be recognized.  We
% should be clear in identifying survey bias; geo bias in claims data; and
% backfill and holiday corrections in claims data.

% Finish with list of iniatives our API can power (like we have now, but
% compress it quite a lot.)

% Signals in the COVIDcast API indicating amounts of COVID-19 and
% COVID-like illness supplement publicly available case and death data by
% capturing and focusing on different levels of severity and parts of the
% timelines of COVID-19 infections. For example, the survey-based COVID-like
% illness indicators may reflect mild or recently-developed COVID-19 infections
% before or without a visit to a healthcare provider or COVID-19
% test. Additionally, the reporting pipeline for various digital signals can be
% quite short. For example, many survey signals are released the morning after the
% surveys were taken, and describe symptoms experienced just one to two days in
% the past; initial estimates of outpatient visits and inpatient admissions are
% released after a few days, generally within a week. The timeliness of these
% signals may enable users to more quickly detect and predict changes in the
% direction of COVID-19 activity.

% TODO move backfill and correlation discussion here?
%
% Discuss other applications of the data:
% - Early warning (informal: just call a human to look at things when the
%   indicators blow up)
% - Hotspot prediction (formal: build a model to predict a surge)
% - Forecasting (refer out to the paper we're writing in parallel)
% - Nowcasting (i.e., estimating latent infection curve)
% - Reopening criteria (refer to COVID Exit Strategy?)
%
% COVIDcast facilitates several types of real-time applications and retrospective
% analyses.
% \begin{itemize}
% \item
% \emph{Situational awareness dashboards and reports}
% summarize activity levels and outcomes, public behavior, vaccination rates, and
% other relevant statistics.
% %
% Features of the COVIDcast API and \texttt{covidcast} packages that are
% particularly relevant here are the common data format across all signals and the
% ability to easily request data for individual locations and for less-commonly
% used geographical resolutions such as Hospital Referral Regions and metropolitan
% statistical areas, and for newer types of data such as vaccination status.
% %
% \item
% \emph{Early warning} systems flag fast rises in COVID-19 activity indicators;
% formal variants first define what counts as elevated activity or surges and
% perform statistical \emph{hotspot predictions}.
% %
% For individuals and local-level response, signals with sufficient availability
% and high time-wise correlations with an outcome of interest (e.g., official case
% or hospitalization data) are natural choices to monitor.
% %
% For resource allocation across locations or drug trial site selection, natural
% choices would be indicators directly describing the stage of the disease being
% targeted or signals with high geo-wise correlations to these indicators.
% %
% \item
% \emph{Forecasts} directly predict COVID-19 activity levels using point estimates
% or probability distributions.
% %
% These models may operate at fine geographical and temporal resolutions and/or
% draw on a wide variety of data, making a common data format and signals offered
% by county and by date particularly useful.
% %
% Revision tracking is especially useful here, helping to explain anomalies in
% past forecasts and improve future ones.
% %
% \item
% \emph{Nowcasts} estimate current or near-future COVID-19 activity levels or the
% true number of infections.
% %
% For these applications, it is important to have signals with time
% values that relate to when symptoms started or were present, or when a healthcare
% visit occurred, rather than to the time that these values were eventually
% reported, unless there is a fixed reporting delay or auxiliary data regarding the
% distribution of reporting delays.
% %
% For example, time values for survey data in COVIDcast relates to when symptoms
% were occurring, doctor visit data indicates the date of the visit, and inpatient
% data indicates the date of admission.
% %
% This contrasts with nationwide aggregated case and death data, which is
% usually aligned by the date that a case or death was reported and subject to
% variable delays based on testing availability and location-specific reporting pipelines.
% %
% % TODO antibodies, revisions, positivity
% \item
% \emph{Reopening criteria} are informed by COVID-19 activity levels, public
% behavior, and vaccination levels.  Methodologies to develop and evaluate these
% criteria may require such information, as well as additional data on timing of
% public health interventions and on confounding variables, in order to draw valid
% conclusions or to improve estimated effects of potential policies.
% \end{itemize}

% TODO refer out to the paper we're writing in parallel

>>>>

\dropcap{P}ublic health decision makers, healthcare providers, epidemiological 
researchers, employers, institutions, and the general public benefit from
promptly and readily accessible data regarding COVID-19 activity levels,
countermeasures, and pandemic impact.  Real-time indicators of COVID-19 
activity levels, such as statistics on cases, deaths, test positivity, and
hospitalizations, enable reports and interactive dashboard applications for
situational awareness \cite{Dong:2020, NYTimes, USAFacts}, and are essential for
most analyses of the pandemic.  These data are available for locations across
the United States from a number of official sources and independent aggregators
in varied and inconsistent formats.  Different data types and sources vary in
timeliness, based on when measured events occur in the progression of the
disease, testing capabilities, the reporting pipeline, and their publication
schedules.

@@

The Delphi Group works with partner organizations and public data sets to build
a massive database of indicators tracking COVID-19 activity and other relevant
phenomena in the United States, which has been publicly available and
continuously updated since April 2020. Alongside public data on reported cases
and deaths, this database includes several unique data streams, including
indicators extracted from de-identified medical claims data, antigen test
results from a major testing manufacturer, large-scale public surveys that
measure symptoms and public behavior, and indicators based on particular Google
search queries. (We use the terms ``indicator'' and ``signal'' interchangeably.)
We make aggregate signals publicly available, generally at the county level, via
the COVIDcast API \cite{CovidcastAPI}. We store and provide access to all
previous (historical) versions of the signals, a key feature that exposes the
effects of data revisions.  Lastly, we provide R \cite{CovidcastR} and Python
\cite{CovidcastPy} packages to facilitate interaction with the API, and an
online dashboard to visualize the data \cite{CovidcastViz}.

@@

\paragraph{Internet-Based Surveys} Conducted in partnership with Facebook,
Delphi's COVID Trends and Impact Survey receives an average of 50,000 responses
daily, and has received over 20 million responses since April 2020
\cite{DelphiSurvey, Kreuter:2020}. From the surveys, we construct indicators on
symptoms, social distancing, vaccination, and other attitudes and behaviors
related to COVID.

@@

\begin{table*}[t]
  \centering
  \caption{Data sources available in Delphi's COVIDcast API \cite{CovidcastAPI},
    as of date of publication. The first group of data sources are produced by
    Delphi from data not otherwise available publicly (or only available in
    limited form); the second group is mirrored from public sources.}
  \begin{tabular}{>{\raggedright}p{1.2in} p{4.0in} l >{\raggedright\arraybackslash}p{0.5in}}
    \toprule
    \textbf{Data source} & \textbf{Signals available} & \textbf{First date} &
\textbf{Resolution} \\\midrule
    % Exclusive data
    Change Healthcare & Percentage of outpatient visits with COVID diagnostic
codes or codes indicating COVID-like symptoms. Based on de-identified claims
data processed by Change Healthcare. & 2020-02-01 & County$^*$ \\
    Doctor Visits & Percentage of outpatient visits primarily about COVID-like
symptoms, based on de-identified claims data provided by health system
partners. & 2020-02-01 & County$^*$ \\
    Hospital Admissions & Percentage of new hospital admissions with COVID
diagnostic codes, based on de-identified claims data provided by health system
partners. & 2020-02-01 & County$^{**}$\\
    Quidel & Test positivity rates for COVID-19 antigen tests produced by
Quidel. & 2020-05-26 & County$^{**}$ \\
    SafeGraph & Mobility metrics, such as time away from home or visits to bars 
and restaurants, based on cell phone mobility data collected by SafeGraph
\cite{SafeGraphSocial, SafeGraphPatterns}. & 2019-01-01 & County \\
    COVID Trends and Impact Survey & COVID symptoms, social distancing 
behaviors, mental health, economic impact, behavior (e.g., mask-wearing,
vaccination attitudes), and COVID testing signals based on daily surveys
conducted nationally by Delphi through Facebook \cite{DelphiSurvey,
Kreuter:2020}. & 2020-04-06 & County$^{**}$ \\
    \midrule
     % Mirrored data from other sources
    Health \& Human Services & Counts of hospital admissions due to confirmed or 
suspected COVID-19, as reported by the Department of Health \& Human Services. &
2019-12-31 & State \\
    COVID Act Now & COVID-19 testing results, such as positivity rate and number
of tests, compiled by COVID Act Now from CDC reporting. & 2020-03-02 &
County$^*$ \\
    Google Symptoms & Trends in Google search volume for terms related to
anosmia and ageusia (loss of smell or taste), which correlate with COVID
activity, based on data shared by Google \cite{GoogleSymptoms}. & 2020-02-13 &
County$^{***}$ \\
    Cases and Deaths & Confirmed COVID-19 cases and deaths, compiled by JHU CSSE 
\cite{Dong:2020} and by USAFacts \cite{USAFacts}. & 2020-01-22 & County \\
    NCHS Mortality & Weekly totals of deaths broken down by cause, such as
COVID, flu, or pneumonia, compiled by the National Center for Health Statistics
\cite{TODO}. & 2020-01-26 & State \\ 
    \bottomrule
  \end{tabular}
  \addtabletext{$^*$Available at $>60$\% of counties; $^{**}$available at
    20--60\% of counties; $^{***}$available at $<20$\% of counties.  For some
    signals, location availability varies over time, e.g., due to variable
    reporting volume.}
  \label{tab:sources_signals}
\end{table*}

@@

Because each data source reports data in different formats, we must convert each
source to a common format. In this format, each record represents an observation
of one quantity at one time point in one location. Locations are coded
consistently using standard identifiers such as FIPS codes; the sample size and
standard error for each observation is also reported when applicable. Each
signal is reported at the finest geographic resolution its source supports (such
as county or state) and also aggregated to metropolitan statistical areas,
Health and Human Services regions, and hospital referral regions. National
averages are also provided. Crucially, each record is tagged with an
\textit{issue date} referring to when the value was first issued, as described 
below. This allows tracking of revisions made to individual observations, as
each revision is tagged with its own issue date.

When appropriate, additional post-processing (often nontrivial) is applied to
the data. For example, data on visits to doctor's offices is subject to strong
day-of-week effects, and so regression is used to adjust for these
effects. Other indicators are available in raw versions and versions smoothed 
with a 7-day trailing average. All processing is done using open-source code 
written primarily in Python and R, and available publicly at
\url{https://github.com/cmu-delphi/covidcast-indicators/}.

@@

Since April 2020, Delphi has been maintaining and continually improving various
online visualization tools for the COVIDcast indicators \cite{CovidcastViz}.
These tools fetch data directly from the API, and allow for exploration of both 
temporal (e.g., time series graphs) and spatial (e.g., choropleth maps) trends
in the signals, as well as many other aspects, such as correlations, anomalies,
and backfill.  There is also a dedicated dashboard for exploring results from
the COVID Trends and Impact Survey.

@@

Many of the indicators in the COVIDcast API are intended to track COVID
activity. Five indicators in particular have the closest connections to
confirmed cases:

\begin{itemize}
\item Change Healthcare COVID-like illness (CHNG-CLI): The percentage of
  outpatient visits that are primarily about COVID-related symptoms, based on
  de-identified Change Healthcare claims data.
\item Change Healthcare COVID (CHNG-COVID): The percentage of outpatient visits
  with confirmed COVID-19, based on the same claims data.
\item COVID Trends and Impact Survey CLI (CTIS-CLI): The estimated percentage
  of the population with COVID-like illness based on Delphi's surveys of
  Facebook users.
\item COVID Trends and Impact Survey CLI in the community
  (CTIS-CLI-in-community): The estimated percentage of the population who knows
  someone in their local community that is sick, based the same surveys.
\item Quidel test positivity rate (Quidel-TPR): The percentage of positive
  results among Quidel COVID antigen tests.  
\end{itemize}

Figure~\ref{fig:time_trends_national} compares the first three of these signals
to COVID cases in the United States (from JHU CSSE, smoothed with a 7-day
trailing average) over a year of the pandemic (April 15, 2020 to April 15,
2021), illustrating how they track national trends quite well.  Importantly, 
this same relationship persists across multiple resolutions of the data, down to
smaller geographic regions such as states and counties, as shown in the
supplement.  This will also be illustrated in a more detailed correlation 
analysis in the next subsection.

% TODO
% RJT: @AR Note what I claimed we would do above!  I'm thinking we can just
% produce a big batch of case trends over the same time period and for the same
% signals, but for each US state (and territory?), and for the 50 most populous
% counties. AND maybe for this (below national) we should use CLI-in-community
% rather than CLI, from the survey

@@

% TODO 
% RJT: @AR Consider for the supplement.  [Or skip it?  Not all that important 
% but I think it would be nice to do.  Maybe you can delegate this to Logan,
% Aaron, Jingjing, etc.] 
% - Hospitalization trends?
% - Hospitalization correlations?
%  (I see you already have a plot for the first ... but I'm a bit bothered by
%  the fact that the HHS data looks to have such strong weekday effects that we
%  don't attempt to correct for ... perhaps we use do a 7-day rolling avg. Made
%  easy with modeltools)
% If we do so, we should refer to it in the text, this subsection (for time
% trends) and next (for correlations)

@@

To quantify the ability of the signals described above to track trends in COVID
cases, we use the Spearman (rank) correlation and analyze two key correlation
patterns, between each signal and confirmed COVID case rates (cases per 100,000
people):

\begin{enumerate}
\item \textit{Geo-wise correlations} (i.e., on a specific date, do values of the
  signal correlate with case rates across locations?): Formally, let $X_t$ and
  $Y_t$ be vectors of values of a particular signal and case rates, over all
  locations, on date $t$. The geo-wise correlation at time $t$ is defined as
  $\cor(X_t, Y_t)$. This examines whether a signal has the capability to help
  spot locations with high case rates at any given time.

@@

Figure~\ref{fig:geo_wise_correlation} shows the geo-wise correlations achieved
by the five signals and COVID case rates (from USAFacts, smoothed using a 7-day
trailing average), from April 15, 2020 to April 15, 2021. This calculation is
performed over counties with at least 500 cumulative cases by the end of this 
period, and at which all indicators are available (956 counties in total). The
significantly positive correlations suggest that these signals could be useful
in hotspot detection (identifying counties with relatively high COVID activity,
at a given time). Somewhat surprisingly, the survey-based CLI-in-community
signal shows the strongest correlations for much of the time period. This
clearly demonstrates the value of a large-scale survey such as CTIS for tracking
symptoms and case trends, especially when other data is unavailable.

@@

Figure~\ref{fig:time_wise_correlation} summarizes time-wise correlations from 
these five signals over the same time period, and for the same set of
counties. For each signal, we display the set correlations that it achieves in
histogram form (more precisely, using a kernel density estimate). All signals
produce positive correlations in the majority of counties considered (with very
little mass in each estimated density being to the left of zero). The largest
correlations, in bulk, are achieved by the CHNG-COVID signal; the
CTIS-CLI-in-community is a close second, and the CHNG-CLI signal is third.
There are two noteworthy points:

\begin{itemize}
\item This is different from what is observed in
  Figure~\ref{fig:geo_wise_correlation}, where the CTIS-CLI-in-community signal
  achieves clearly the highest correlations for most of the time
  period. However, it is worth emphasizing that time-wise and geo-wise
  correlations are truly measuring different properties of a signal; and the
  claims signals (CHNG-COVID and CHNG-CLI) seem more appropriate for
  temporal---rather than spatial---comparisons.  We revisit this point in the
  discussion.

\item It is still quite impressive (and surprising) that the
  CTIS-CLI-in-community signal, based on people reporting on the symptoms of
  others around them, can achieve  nearly as strong time-wise correlations to
  confirmed cases as can a signal that is based on picking up the occurrence of
  a confirmed case passing through the outpatient system.
\end{itemize}

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=\columnwidth]{fig/bexar_compare.pdf}
  \caption{Reported cases per day in Bexar County, Texas during the summer of
    2020. Only July 16, 4,810 backlogged cases were reported, though they
    actually occurred over the preceding two weeks (this shows up as a prolonged
    spike in the left panel due to the 7-day trailing averaging applied to the
    case counts). Daily CTIS estimates of CLI-in-community showed more stable
    underlying trends.}
  \label{fig:bexar_compare}
\end{figure}

@@

As a further example, Figure~\ref{fig:bexar_compare} displays data from Bexar
County, Texas (which contains San Antonio) during July 2020. On July 16, 2020,
4,810 backlogged cases were reported after reporting problems prevented them
from being reported over the past two weeks \cite{Palacios:2021}, resulting in a
clearly visible spike in the left-hand panel of Figure~\ref{fig:bexar_compare}
(case data from JHU CSSE, smoothed using a 7-day trailing average).  Meanwhile,
Delphi's COVID Trends and Impact Survey averaged around 350 responses per day in
Bexar County over the same time period, and was able to estimate the fraction of
the population who know someone in their community with COVID-Like Illness
(CLI). As we can see in the right-hand panel of Figure~\ref{fig:bexar_compare},
this signal was not affected by Bexar County's reporting problems and, as shown
in the last subsection, it is (in general) highly correlated with case rates,
providing an alternate stream of data about COVID activity unaffected by
backlogs. Similar reporting problems have occurred in many jurisdictions across
the United States, making it valuable to cross-check against external sources
not part of the same reporting systems.

@@

% TODO

@@

% RJT: @AR Since we already stuff on claims, can we do: mobility, masks, vaccine
% uptake or hesitancy, distancing behaviors, how worried are you?
% How about this (in each case, each point in said plot is a county's signal
% values averaged over some period)
% - Bivariate plot of mask use versus how worried are you
% - Bivariate plot of of how worried are you vs case rates
%- Bivariate plot of of how worried are you vs hospitalizations
% I would be curious to see all three.  And perhaps color code the points by
% region to highlight regional trends

@@

The COVIDcast API provides open access to real-time and geographically-detailed
indicators that support and enhance standard public health reporting streams in 
several ways.

First, several signals in the API closely track COVID activity (over both time
and space); yet they are derived from different data streams (e.g., surveys,
medical insurance claims, medical devices), and are therefore not subject to the
same sources of error as public health reporting data.  This can be important
both for downstream modeling purposes (enhancing the accuracy in nowcasting and
forecasting models; see our companion paper on forecasting) and also for
robustness (providing sanity checks when faced with potential anomalies in
standard surveillance streams).

Second, the API features many other signals that are relevant to understanding
aspects of the pandemic and its effects on the United States population that are
not found in traditional public health streams, e.g., data on mobility patterns,
internet search trends, mask wearing, and vaccine hesitancy, to name just a few
(the last two are derived from the COVID Trends and Impact Survey; see our
companion paper on this survey for a detailed view of its features and
capabilities).

Third, the underlying database tracks all revisions made to the data, allowing
us to query the API to learn ``what was known when'', critical for understanding
the behavior (and potential pitfalls) of real-time surveillance signals.

Finally, we emphasize that unifying many relevant signals into a single common
format, with comprehensive revision tracking, is a valuable goal in it of
itself.  No one data source is perfect; each is subject to error.  Bringing
numerous signals from different data streams all together, and making access as
convenient as possible for public health researchers, policy makers, modelers,
etc., is a strong testament to our belief in ``the whole being greater than the
sum of its parts''.

There are a number of open questions, and challenges that remain.  
% TODO:
% RJT: we should write quick bullet points on 1. survey bias, 2. geo bias in
% claims data, 3. backfill and holiday corrections in claims data. 
% @AR: can you give this a stab?  You can also delegate (Aaron e.g. should be
% able to write something on the claims stuff.)  Let's keep it succint!
